---
title: "SARIMA Models with Transformations and Seasonal Differencing"
author: "Evan L. Ray"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{SARIMA Utilities}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

This vignette demonstrates the functionality provided by the `sarimaTD`
package using influenza data available from the cdcfluview package.  The goal of
this package is to provide a set of functions for estimating and generating
predictions from a SARIMA model for approximately seasonal infectious disease
time series data that comes as close as possible to working "out of the box"
with minimal adjustments required.

## Dengue Fever Data

We will work with the `san_juan_dengue` data provided in this package.  Note
that these data are the "Training Data" available for San Juan, Puerto Rico from
NOAA here: http://dengueforecasting.noaa.gov/

```{r}
library(sarimaTD)
library(dplyr)
library(ggplot2)

ggplot(data = san_juan_dengue,
    mapping = aes(x = week_start_date, y = total_cases)) +
  geom_line()
```

Let's fit models using the data up through the 2007/2008 season and make
predictions for the 2008/2009 season.

## SARIMA Estimation

Let's fit a few variations on models using different transformations, with and
without an initial seasonal differencing.  Cite Hyndman, who recommends a
preliminary seasonal differencing somewhere...

Note that stationary = TRUE for auto.arima forces all differencing = 0 because
resulting model is non-stationary if differencing is used.

```{r}
train_indices <- san_juan_dengue$season %in% paste0(1990:2004, "/", 1991:2005)

sarima_fit_bc_transform <- fit_sarima(
  y = san_juan_dengue$total_cases[train_indices],
  ts_frequency = 52,
  transformation = "box-cox",
  seasonal_difference = TRUE)

#sarima_fit_log_transform <- fit_sarima(
#  y = san_juan_dengue$total_cases[train_indices] + 1, # there are some observations of 0 cases
#  ts_frequency = 52,
#  transformation = "log",
#  seasonal_difference = TRUE)

#sarima_fit_no_transform <- fit_sarima(
#  y = san_juan_dengue$total_cases[train_indices],
#  ts_frequency = 52,
#  transformation = "none",
#  seasonal_difference = TRUE)
```

## SARIMA Prediction

```{r}
eval_indices <- san_juan_dengue$season %in% paste0(1990:2005, "/", 1991:2006)

sampled_trajectories_bc_transform <-
  simulate(
    object = sarima_fit_bc_transform,
    nsim = 1000,
    seed = 1,
    newdata = san_juan_dengue$total_cases[eval_indices],
    h = 52
  )

plot_indices <- san_juan_dengue$season %in% paste0(2003:2006, "/", 2004:2007)

preds_df <- san_juan_dengue %>%
  filter(season == "2006/2007") %>%
  mutate(
    pred_total_cases = apply(sampled_trajectories_bc_transform, 2, mean),
    pred_95_lb = apply(sampled_trajectories_bc_transform, 2, quantile, probs = 0.025),
    pred_95_ub = apply(sampled_trajectories_bc_transform, 2, quantile, probs = 0.975),
    pred_80_lb = apply(sampled_trajectories_bc_transform, 2, quantile, probs = 0.05),
    pred_80_ub = apply(sampled_trajectories_bc_transform, 2, quantile, probs = 0.95),
    pred_50_lb = apply(sampled_trajectories_bc_transform, 2, quantile, probs = 0.25),
    pred_50_ub = apply(sampled_trajectories_bc_transform, 2, quantile, probs = 0.75),
  )

ggplot() +
  geom_ribbon(
    mapping = aes(x = week_start_date, ymin = pred_95_lb, ymax = pred_95_ub),
    fill = "cornflowerblue",
    alpha = 0.2,
    data = preds_df) +
  geom_ribbon(
    mapping = aes(x = week_start_date, ymin = pred_80_lb, ymax = pred_80_ub),
    fill = "cornflowerblue",
    alpha = 0.2,
    data = preds_df) +
  geom_ribbon(
    mapping = aes(x = week_start_date, ymin = pred_50_lb, ymax = pred_50_ub),
    fill = "cornflowerblue",
    alpha = 0.2,
    data = preds_df) +
  geom_line(
    mapping = aes(x = week_start_date, y = pred_total_cases),
    color = "cornflowerblue",
    data = preds_df) +
  geom_line(mapping = aes(x = week_start_date, y = total_cases),
    data = san_juan_dengue[plot_indices, , drop = FALSE])
```
