---
title: "SARIMA Models with Transformations and Seasonal Differencing via ForecastFramework"
author: "Evan L. Ray"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{sarimaTD ForecastFramework interface}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

The ForecastFramework package specifies a standardized interface for forecasting
methods.  The sarimaTD package provides an R6 class, `sarimaTD_FF`, which can
be used with the ForecastFramework package.  This vignette demonstrates the use
of `sarimaTD_FF`.

## Dengue Fever Data

We will work with the `san_juan_dengue` data provided in this package.  Note
that these data are the "Training Data" available for San Juan, Puerto Rico from
NOAA here: http://dengueforecasting.noaa.gov/

```{r}
library(sarimaTD)
library(ForecastFramework)
library(dplyr)
library(ggplot2)

ggplot(data = san_juan_dengue,
        mapping = aes(x = week_start_date, y = total_cases)) +
    geom_line()
```

Let's fit models using the data up through the 2007/2008 season and make
predictions for the 2008/2009 season.

## SARIMA Estimation

```{r}
# get a train/test split
san_juan_dengue <- san_juan_dengue %>%
    mutate(location = "San Juan, PR")
train_seasons <- paste0(1990:2004, "/", 1991:2005)
train_indices <- which(san_juan_dengue$season %in% train_seasons)
train_data <- san_juan_dengue %>%
    slice(train_indices)
test_data <- san_juan_dengue %>%
    slice(-train_indices)

# get data into format for ForecastFramework model
train_data_FF <- ObservationList$new(train_data)
train_data_FF$formArray('location',
    'week_start_date',
    val = 'total_cases',
    dimData = list(list('location'), list('week_start_date')))

test_data_FF <- ObservationList$new(test_data)
test_data_FF$formArray('location',
    'week_start_date',
    val = 'total_cases',
    dimData = list(list('location'), list('week_start_date')))

# define sarimaTD_FF object
sarima_model <- sarimaTD_FF$new(
    nsim = 1000,
    frequency = 52,
    transformation = "box-cox",
    seasonal_difference = TRUE)

# get the model fit to the training data
sarima_model$fit(train_data_FF)
```

## SARIMA forecasts

Now we generate some forecasts from the model and plot them.

```{r}
forecast_FF <- sarima_model$forecast(steps = 52)

plot_indices <- san_juan_dengue$season %in% paste0(2003:2006, "/", 2004:2007)

preds_df <- san_juan_dengue %>%
    filter(season == "2006/2007") %>%
    mutate(
        pred_total_cases = as.vector(forecast_FF$median(na.rm=TRUE)$mat),
        pred_95_lb = as.vector(forecast_FF$quantile(0.025,na.rm=TRUE)$mat),
        pred_95_ub = as.vector(forecast_FF$quantile(0.975,na.rm=TRUE)$mat),
        pred_80_lb = as.vector(forecast_FF$quantile(0.05,na.rm=TRUE)$mat),
        pred_80_ub = as.vector(forecast_FF$quantile(0.95,na.rm=TRUE)$mat),
        pred_50_lb = as.vector(forecast_FF$quantile(0.25,na.rm=TRUE)$mat),
        pred_50_ub = as.vector(forecast_FF$quantile(0.75,na.rm=TRUE)$mat)
    )

ggplot() +
    geom_ribbon(
        mapping = aes(x = week_start_date, ymin = pred_95_lb, ymax = pred_95_ub),
        fill = "cornflowerblue",
        alpha = 0.2,
        data = preds_df) +
    geom_ribbon(
        mapping = aes(x = week_start_date, ymin = pred_80_lb, ymax = pred_80_ub),
        fill = "cornflowerblue",
        alpha = 0.2,
        data = preds_df) +
    geom_ribbon(
        mapping = aes(x = week_start_date, ymin = pred_50_lb, ymax = pred_50_ub),
        fill = "cornflowerblue",
        alpha = 0.2,
        data = preds_df) +
    geom_line(
        mapping = aes(x = week_start_date, y = pred_total_cases),
        color = "cornflowerblue",
        data = preds_df) +
    geom_line(mapping = aes(x = week_start_date, y = total_cases),
        data = san_juan_dengue[plot_indices, , drop = FALSE])
```
